---
title: "Practical Machine Learning Assignment"
author: "Kiran"
date: "Sunday, March 22, 2015"
output: html_document
---

###Loading the data:
The data was downloaded from the following urls:
training set: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
test set: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

Data has been dowloaded and stored on the C drive.
75% of the data is used for training and the rest for validation.
```{r}
TriningData <- read.csv('C:/pml-training.csv');
library(caret);
InTrain <- createDataPartition(y = TriningData$classe, p=0.75, list = F);
training <- TriningData[InTrain,];
testing <- TriningData[-InTrain,];
```

###Removing variables which are not features
we can see there are many sensor output which mostly is empty. Removing all such variables.
```{r, results='hide'}
head(training);
summary(training);
training <- training[,c("roll_belt", "pitch_belt", "yaw_belt", "total_accel_belt", "gyros_belt_x", "gyros_belt_y", "gyros_belt_z", "accel_belt_x", "accel_belt_y", "accel_belt_z", "magnet_belt_x", "magnet_belt_y", "magnet_belt_z", "roll_arm", "pitch_arm", "yaw_arm", "total_accel_arm", "gyros_arm_x", "gyros_arm_y", "gyros_arm_z", "accel_arm_x", "accel_arm_y", "accel_arm_z", "magnet_arm_x", "magnet_arm_y", "magnet_arm_z", "roll_dumbbell", "pitch_dumbbell", "yaw_dumbbell", "total_accel_dumbbell", "gyros_dumbbell_x", "gyros_dumbbell_y", "gyros_dumbbell_z", "accel_dumbbell_x", "accel_dumbbell_y", "accel_dumbbell_z", "magnet_dumbbell_x", "magnet_dumbbell_y", "magnet_dumbbell_z", "roll_forearm", "pitch_forearm", "yaw_forearm", "total_accel_forearm", "gyros_forearm_x", "gyros_forearm_y", "gyros_forearm_z", "accel_forearm_x", "accel_forearm_y", "accel_forearm_z", "magnet_forearm_x", "magnet_forearm_y", "magnet_forearm_z", "classe")]
```

###Removing variables which are highly correlated:
Some of the variables would be highly correlated. Removing some of the variables which will not have much effect.
```{r}
##Calculates Pairwise correlation
pairwiseCor <- function(dataframe){
  pairs <- combn(names(dataframe), 2, simplify=FALSE)
  df <- data.frame(Vairable1=rep(0,length(pairs)), Variable2=rep(0,length(pairs)), 
                   AbsCor=rep(0,length(pairs)), Cor=rep(0,length(pairs)))
  for(i in 1:length(pairs)){
    df[i,1] <- pairs[[i]][1]
    df[i,2] <- pairs[[i]][2]
    df[i,3] <- round(abs(cor(dataframe[,pairs[[i]][1]], dataframe[,pairs[[i]][2]])),4)
    df[i,4] <- round(cor(dataframe[,pairs[[i]][1]], dataframe[,pairs[[i]][2]]),4)
  }
  pairwiseCorDF <- df
  pairwiseCorDF <- pairwiseCorDF[order(pairwiseCorDF$AbsCor, decreasing=TRUE),]
  row.names(pairwiseCorDF) <- 1:length(pairs)
  pairwiseCorDF <<- pairwiseCorDF
  pairwiseCorDF
  }


training.cor <- pairwiseCor(training[, names(training) != 'classe'])
training.cor[training.cor$AbsCor>0.8,];

##Removing highly correlated columns
training <- training[,c("yaw_belt", "gyros_belt_x", "gyros_belt_y", "gyros_belt_z", "accel_belt_x", "accel_belt_y", "accel_belt_z", "magnet_belt_x", "magnet_belt_y", "magnet_belt_z", "roll_arm", "pitch_arm", "yaw_arm", "total_accel_arm", "gyros_arm_x", "gyros_arm_y", "gyros_arm_z", "accel_arm_x", "accel_arm_y", "accel_arm_z", "magnet_arm_x", "magnet_arm_y", "magnet_arm_z", "roll_dumbbell", "pitch_dumbbell", "yaw_dumbbell", "total_accel_dumbbell", "gyros_dumbbell_x", "gyros_dumbbell_y", "accel_dumbbell_x", "accel_dumbbell_y", "accel_dumbbell_z", "magnet_dumbbell_x", "magnet_dumbbell_y", "magnet_dumbbell_z", "roll_forearm", "pitch_forearm", "yaw_forearm", "total_accel_forearm", "gyros_forearm_x", "gyros_forearm_y", "accel_forearm_x", "accel_forearm_y", "accel_forearm_z", "magnet_forearm_x", "magnet_forearm_y", "magnet_forearm_z", "classe")]
```


Building the prediction model:
I am using the prediction method 'boosting with trees' because:
1. The data is non-linear.
2. The different sensor data may be week predictors and this method will combine them to give strong predictors.
```{r cache=TRUE}
set.seed(1000);
predictModel <- train(classe ~ ., method = 'gbm', data = training, verbose = F);
```

###Predicting the Training set:
```{r}
prediction <- predict(predictModel,training);
table(prediction, prediction == training$classe);
```

###Predicting the Testing set: Out of sample error:
```{r}
prediction <- predict(predictModel,testing);
table(prediction, prediction == testing$classe);
```
As expected the error ratio is higher in the testing set.
